### 2009-09-02 doc/tech_doc/janitor/tex_introduction/introduction.tex
Adding Janitor documentation from janitor branch.

Janitor is a service for the automated installation of runtime environment for grid computing elements.
It is addressed transparently via the A-REX service for job submission but can also be used as a standalone tool.

@@

A core problem remains to distribute a locally working solution, the Know-How, quickly across all
contributing sites, i.e., without manual interference. Every scientific discipline has its respective own set of
technologies for the distribution of work load. E.g. research in bioinformatics requires access to so many
different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
a commodity, with all their intrinsic problems as there are bottlenecks and restrictions of repeated access.
The EU project KnowARC\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges
with the here presented work extends the NorduGrid's
Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an infrastructure for the
automated installation of software packages.

An automation of the software installation, referred to as dynamic Runtime Environments, seems the only
approach to use the computational grid to its full potential. Components of workflows shall be spawned
as jobs in a computational grid using dynamic Runtime Environments rather than accessing a web service
at one particular machine that is shared amongst all users.
The grid introduces an extra level of parallelism that web services cannot provide. The demands for short
response times and the heterogeneous education of site-administrators on a grid demand an automatism for
the installation of software and databases without manual interference~\cite{BAYER_2007}.

@@

This document first dedicates a chapter on how to set-up the Janitor locally. It is followed by a chapter that gives furhter instructions on how to 
use the Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
which is basically covering the method how to prepare new runtime environments. Deeper insights on the design of the Janitor will 
be given by the forth chapter. The document ends with an outlook to anticipated future developments and opportunities.

\section*{Abbreviations}

\begin{tabular}{l@{--}p{150mm}}
RTE&Runtime Environment\\
dRTE&dynamic Runtime Environment\\
RDF&Resource Description Framework (supporting the RTE Catalog)\\
\end{tabular}

>>>>

Janitor is a service for automated runtime environment installation for grid computing elements. It is integrated in the A-REX service used for job submission but also can be used as a standalone tool.

@@

A core problem remains to distribute a locally working solution, the Know-How, quickly across all contributing sites, i. e., without manual interference. Every scientific discipline has its respective own set of
technologies for the distribution of work load. E. g. research in bioinformatics requires access to so many
different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
a commodity, with all the problems with respect to bottlenecks and restrictions of repeated access.
The EU project KnowARC1\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges with the here presented work extends the NorduGrid?s
Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an infrastructure for the
automated installation of software packages.

An automation of the software installation, referred to as dynamic Runtime Environments, seems the only
approach to use the computational grid to its full potential. Components of workflows shall be spawned
as jobs in a computational grid using dynamic Runtime Environments rather than as shared web services.
The grid introduces an extra level of parallelism that web services cannot provide. The required short
response times and the heterogeneous education of site-administrators on a grid demand an automatism for
the installation of software and databases without manual interference~\cite{BAYER_2007}.

@@

This document will start with a chapter of how to set up janitor locally. The following chapter will give an instruction on how to 
use Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
which is basically covering the method how to prepare new runtime environments. A deeper knowledge about the design of Janitor will 
be given by the subsequent forth chapter. In the fifth chapter, an outlook to future developments will given.

### 2009-09-02 doc/tech_doc/janitor/tex_maintenance/maintenance.tex
Adding Janitor documentation from janitor branch.

This chapter explains how to maintain the {\em Catalog} and the {\em Janitor} itself. 
It gives detailed instructions how to create new packages for the Janitor.
To administrate the catalog, it is strongly recommended (but not ultimately required) to use the ontology editor Prot\`eg\`e,
as explained in the next section.
In the last section, a typical use case in maintaining the Janitor will be shown.

@@

The Catalog describes runtime environments and is either served through a web server or is distributed together with the Janitor as a regular file. 
It is specifed by a (Resource Description Framework) RDF file assigned to 
Janitor using the tag \texttt{catalog} within the configuration file (see~\ref{sec:janitor_configuration}). 
The format of the RDF file is defined by a RDF schema file \texttt{knowarc.rdfs}
which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.\\

In order to adminstrate the catalog, the ontology editor Prot\`eg\`e should be used. Figure~\ref{fig:protege_example} shows the
editor while the MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has been selected.
\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{images/protege_JASPAR.png}
    \mycaption{Example of a RDF catalog file as displayed in the program Prot\`eg\`e.}{}
    \label{fig:protege_example}
  \end{center}
\end{figure}
On the left side of the editor the class browser is placed. Three main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and \texttt{Package}.
The \texttt{Metapackage} is a general platform independent description of a \texttt{Package}. 
It has one or more instances of the class \texttt{Package} and is
described by the subclasses of \texttt{Note}. The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the Debian release a \texttt{Package} refers to
(i.e. here etch or sid), i.e. the name of a common installation or a virtual image.
The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
they can be found more easily.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
classes are interacting with each other the tables~\ref{tab:main_class_meta}, \ref{tab:main_class_base}, \ref{tab:main_class_tag},
\ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.

@@

The dynamic Runtime Environments stored in the Catalog are presented on the aforementioned dedicated web
page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
RDF syntax and its automated transformation to
HTML. The latter mimics the traditional site describing Runtime Environments in the Runtime Environment
Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with an eventual transition to the new system. 
That page collects descriptions for Runtime Environments to encourage human site administrators to install these.
This HTML page listing the manually or automatically installable REs is prepared by the script web/list.pl.
This script is meant to be run by a mod\_perl-enabled Apache. The script itself does not contribute to the 
core functionality of the Janitor.
In the first lines of the script some variables specific to the site are set. To configure the
script these have to be changed~\cite[p. 9]{BAYER_2007}.

@@

This section describes how to add new packages to the Catalog. In the current implementation, only tar based packages are processed by the Janitor.
Within the example they are assigned to be used together with Debian Etch. This limitation is only literal. 
There is no limitiation to newer Debian distributions.

@@

At the time of writing, only the tape archive (tar) file format is accepted for dynamic Runtime Environment installation, 
a well accepted file format throughout the UNIX community. The reflects the traditional manual approach towards RTE in ARC,
for which one directory is made available to all compute nodes. This section explains the inner structure of the tar files. 
Subdirectories are visualised in Figure~\ref{fig:tar_folder}.

@@

The tar file contains two directories, named {\tt control} and {\tt data}. Software is stored in the latter subdirectory,
while the files formally specifying how to deal with such packages are stored in the prior. Upon installation,
the content of the data directory is extracted to some directory \textdollar BAR. 
After this unpacking of the tar file, the Janitor executes the install script provided in the control directory. 
It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install} just
before the tar-package is removed. In most cases \texttt{control/remove} will be empty.
Finally, the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. After
installing the package, the Janitor changes all occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.
Once the tar file was prepared, it must its entry to a RTE Catalogue~\cite[p. 10]{BAYER_2007}.

From such Catalogues, the Janitor finds all information to install packages that possibly have never been installed
on the site before. The offers of RTEs in a catalogue are cross-checked against the local infrastructure and a subset
of the available packages will be accepted as "installable". This list of installable RTEs is forwarded to the
grid information system.

@@

In order to have an impression how the tar files are created, several
prototypes are provided at \url{http://dre.knowarc.eu}.

@@

WEKA_JAR="weka-3-4-8a/weka.jar"
case "\$1" in
	0)	# Just before job submission
		# none
	;;
	1)	# Just before job execution
		# Initialize the java environment
		CLASSPATH="%BASEDIR%/$WEKA_JAR:$CLASSPATH"
		export CLASSPATH
	;;
	2)	# After job termination
		# none
	;;
	*)
		return 1
	;;
esac	
\end{verbatim}
The remove script, which will be executed right before WEKA is deinstalled, is empty. Janitor will delete the directory, so there
is nothing more to be done.\\

%\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
%provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
%dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
%thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
%projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\

To address the concerns of the physicists using ARC, a dynamic runtime environment for the
ATLAS software suite was prepared. It extends prior work on an automated installation that is available at
\href{http://guts.uio.no/atlas/12.0.6/}{http://guts.uio.no/atlas/12.0.6/}. The preparation comprised the following steps:
\begin{itemize}
    \item The file system path specifications in the automated installation scripts were modified using the Janitor
       path variables.
    \item A tarball was prepared containing a directory structure as illustrated in Figure~\ref{fig:tar_folder}. The data directory
       was empty, since the automatic installation script downloads the software from a remote server.
    \item An entry was added to the Catalog file.
\end{itemize}

What sets High Energy Physics software apart is it's sheer size. The package in question takes up more than
5 GB. This was a test illustrating the feasibility of using dynamic REs in High Energy Physics. The application
of the dynamic RE for ATLAS needs to wait for the planned web service extension of the Catalog. With such a
service, e.g. a software manager of a big experiment will be able to deploy software packages on production
sites simply by creating a tarball and adding an entry to the Catalog.

%\subsection{Debian Etch}
%
%These kind of packages are not yet supported.
%
%\subsection{Debian Sid}
%
%These kind of packages are not yet supported.
%
%\subsection{Adminstrating the Catalog}

>>>>

This chapter explains how to maintain the catalog and Janitor itself. 
In order to administrate the catalog, it is absolutly recommened (but not required) to use the ontology editor Prot\`eg\`e. The first
section will explain how this is done.
The second section a detailed explanation how to create new packages for Janitor will be given.
In the last section typical use case in maintaining Janitor will be listed.

@@

The Catalog describes runtime environments and is either served through a web server or distributed together with the Janitor. 
It is specifed by a (Resource Description Framework) RDF file assigned to 
Janitor using the tag \texttt{catlog} within the configuration file (see~\ref{sec:janitor_configuration}). 
The format of the RDF file is defined by a RDF schema file \texttt{knowarc.rdfs}
which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.\\

In order to adminstrate the catalog, the ontology editor Prot\`eg\`e should be used. Figure~\ref{fig:protege_example} shows the
editor while the MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has been selected.
\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{images/protege_JASPAR.png}
    \mycaption{Example of a RDF catalog file as displayed in the program Prot\`eg\`e.}{}
    \label{fig:protege_example}
  \end{center}
\end{figure}
On the left side of the editor the class browser is placed. Three main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and \texttt{Package}.
The \texttt{Metapackage} is a general platform independent description of a \texttt{Package}. 
It has one or more instances of the class \texttt{Package} and is
described by the subclasses of \texttt{Note}. The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the Debian release a \texttt{Package} refers to
(i.e. here etch or sid). The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
they can be easier found.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
classes are interacting with each other the Tables~\ref{tab:main_class_meta},\ref{tab:main_class_base}, \ref{tab:main_class_tag},
\ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.

@@

The dynamic Runtime Environments stored in the catalog are presented on the aforementioned dedicated web
page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
RDF syntax and an automated transformation to
HTML. The latter mimics the traditional site describing Runtime Environments in the Runtime Environment
Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with an eventual transition to the new system. 
That page collects descriptions for Runtime Environments to encourage human site administrators to install these.
This html page listing the manually or automatically installable RTEs is prepared by the script web/list.pl.
This script is meant to be run by a mod-perl enabled Apache. The script itself is only loosely integrated
within the Janitor. In the first lines of the script some variables specific to the site are set. To configure the
script these have to be changed~\cite[p. 9]{BAYER_2007}.

@@

This section describes how to add new packages to the catalog. Currently only tar based packages are processable by Janitor.
Within the example they are assigned to be used together with Debian Etch. This limitation is only literal. 
There should be no limitiation
to newer Debian distributions.

@@

At the time of writing, only the tape archive (tar) file format is accepted for dynamic runtime environment installation, 
a well accepted file format throughout the UNIX community. This section explains the inner structure of the tar files. 
Subdirectories are visualised in Figure~\ref{fig:tar_folder}.

@@

The tar file contains two directories named control and data. Software is stored in the latter subdirectory,
while the files formally specifying how to deal with such packages are stored in the prior. Upon installation,
the content of the data directory is extracted to some directory \textdollar BAR. 
After this unpacking of the tar file, the Janitor executes the install skript provided in the control directory. 
It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install} just
before the tar-package is removed. In most cases \texttt{control/remove} will be empty.
Finally, the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. After
installing the package, the Janitor changes all occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.
To be offered to computing elements for an installation, the such prepared runtime environment must be
announced to a Catalog to which the Janitor on the computing element subscribes~\cite[p. 10]{BAYER_2007}.

@@

In order to have an impression how the tar files are created, several prototypes are provided at \task{URL required!!}.:\\

@@

WEKA_JAR="weka-3-4-8a/weka.jar"
case "\$1" in
	0)	# Just before job submission
		# none
	;;
	1)	# Just before job execution
		# Initialize the java environment
		CLASSPATH="%BASEDIR%/$WEKA_JAR:$CLASSPATH"
		export CLASSPATH
	;;
	2)	# After job termination
		# none
	;;
	*)
		return 1
	;;
esac	
\end{verbatim}
The remove script, which will be executed right before WEKA is deinstalled, is empty. Janitor will delete the directory, so there
is nothing to be done.\\

\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\

To address the concerns of the physicists using ARC, a dynamic runtime environment for the
ATLAS software suite was prepared. It extends prior work on an automated installation that is available at
\href{http://guts.uio.no/atlas/12.0.6/}{http://guts.uio.no/atlas/12.0.6/}. The preparation comprised the following steps:
\begin{itemize}
    \item The file system path specifications in the automated installation scripts were modified using the Janitor
       path variables.
    \item A tarball was prepared containing a directory structure as illustrated in Figure~\ref{fig:tar_folder}. The data directory
       was empty, since the automatic installation script downloads the software from a remote server.
    \item An entry was added to the Catalog file.
\end{itemize}
What sets High Energy Physics software apart is it's sheer size. The package in question takes up more than
5 GB. This was a test illustrating the feasibility of using DREs in High Energy Physics. The application
of the DRE for ATLAS needs to wait for the planned web service extension of the Catalog. With such a
service, e.g. a software manager of a big experiment will be able to deploy software packages on production
sites simply by creating a tarball and adding an entry to the Catalog.

\subsection{Debian Etch}

These kind of packages are not yet supported.

\subsection{Debian Sid}

These kind of packages are not yet supported.

\subsection{Adminstrating the Catalog}

### 2009-09-02 doc/tech_doc/janitor/tex_titlepage/titlepage.tex
Adding Janitor documentation from janitor branch.

%%%%---- The title ----
{\centering \textsc{\Large Dynamic Runtime Environments with Janitor}
\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to glodek@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
\Large \par}
\vspace*{0.5cm}

@@

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Michael Glodek\footnote{glodek@inb.uni-luebeck.de}, Daniel Bayer,  Steffen M\"oller\footnote{moeller@inb.uni-luebeck.de} \par}

>>>>

%%%%---- The title ----
{\centering \textsc{\Large Dynamic Runtime Environment Installation with Janitor}
\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to glodek@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
\Large \par}
\vspace*{0.5cm}

@@

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Michael Glodek\footnote{glodek@inb.uni-luebeck.de}, Daniel Bayer,  Steffen M\"oller \par}
